---
title: "ETC5521 Assignment 1"
subtitle: "Amusement Park injuries"
team: "bilby"
author:
  - "Yuheng Cui"
  - [FILL]
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2
---

[This assignment is for ETC5521 Assignment 1 by Team `r rmarkdown::metadata$team` comprising of `r knitr::combine_words(rmarkdown::metadata$author)`.]{style="color:#006DAE;"}

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.align = "center")
library(dplyr)
library(visdat)
library(readr)
```

```{r rawData, include = FALSE}
tx_injuries <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-10/tx_injuries.csv")


safer_parks <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-10/saferparks.csv")
```

# Introduction and motivation

intro: ...

motivation: study the injuries record, find out the risky groups. In the future, amusement parks can pay more attention on those groups and entertaining equipments. Consequently, reduce the injuries occur in amusement parks and increase the well beings of society.

# Data description

The datasets are downloaded from [here](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-09-10) from GitHub. The origin datasets are from [data.world](https://data.world/amillerbernd/texas-amusement-park-accidents) and [Safer Parks database](https://ridesdatabase.org/saferparks/data/).

The source of data.world dataset is Texas Department of Insurance. This dataset recoreded all incidents at Texas amusement parks from 1st Feb 2013 to 1st Feb 2017. has `r dim(tx_injuries)[1]` observations and `r dim(tx_injuries)[2]` variables. Fig \@ref(fig:quality-of-tx) shows the dataset has too many missing values in **other** variable and many missing values in **serial_no** variable. Good news is these two variables has little effects on our analysis, thus, we can remove these two variables. **Age** should be converted to numeric variable and **injury_date** should be date variable.

```{r quality-of-tx, fig.cap = "Texas amusement parks dataset"}
vis_dat(tx_injuries)
```


The source of SaferParks dataset is The U.S. Consumer Product Safety Commission (CPSC). It has `r dim(safer_parks)[1]` observations and `r dim(safer_parks)[2]` variables. Fig \@ref(fig:quality-of-safer) shows 4 variables have too many missing values to be analysed, therefore, we ought to ignore them. **acc_date** should be converted to date variable. Although **manufacturer** has many missing values as well, we decided to keep it because it may be insightful.

```{r quality-of-safer, fig.cap = "Saferparks dataset"}
vis_dat(safer_parks)
```

# Analysis and findings

[FILL] Should include at least one plot or numerical summary for each of your questions, that helps the reader arrive at an answer. You should also write paragraphs describing the methods, summaries and findings. 

# References
